---
title: "Hands-on_Ex10&11"
author: "Ryan Peh"
date: "October 27, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
---

## **1 Overview**

**Geographically Weighted Regression** (GWR) is a spatial statistical method that accounts for non-stationary variables, such as climate, demographic factors, and physical environment characteristics. GWR models the local relationships between these independent variables and a dependent variable, which is the outcome of interest. In this hands-on exercise, we will construct hedonic pricing models using GWR techniques. The dependent variable in this context is the resale prices of condominiums in 2015, while the independent variables are categorized into structural and locational factors.

## **2 Data**

Two data sets will be used in this model building exercise, they are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)

-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

## **3 Packages**

Before we get started, it is important for us to install the necessary R packages into R and launch these R packages into R environment.

The R packages needed for this exercise are as follows:

-   R package for building OLS and performing diagnostics tests

    -   [**olsrr**](https://olsrr.rsquaredacademy.com/)

-   R package for calibrating geographical weighted family of models

    -   [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/index.html)

-   R package for multivariate data visualisation and analysis

    -   [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)

-   Spatial data handling

    -   **sf**

-   Attribute data handling

    -   **tidyverse**, especially **readr**, **ggplot2** and **dplyr**

-   Choropleth mapping

    -   **tmap**

The code chunks below installs and launches these R packages into R environment.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

::: callout-note
GWmodel package provides a collection of localised spatial statistical methods, namely: GW summary statistics, GW principal components analysis, GW discriminant analysis and various forms of GW regression; some of which are provided in basic and robust (outlier resistant) forms. Commonly, outputs or parameters of the GWmodel are mapped to provide a useful exploratory tool, which can often precede (and direct) a more traditional or sophisticated statistical analysis.
:::

## **4 Spatial Data Wrangling**

### **4.1 Importing geospatial data**

The geospatial data used in this hands-on exercise is called MP14_SUBZONE_WEB_PL. It is in ESRI shapefile format. The shapefile consists of URA Master Plan 2014’s planning subzone boundaries. Polygon features are used to represent these geographic boundaries. The GIS data is in svy21 projected coordinates systems.

The code chunk below is used to import *MP_SUBZONE_WEB_PL* shapefile by using `st_read()` of **sf** packages.

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

The report above shows that the R object used to contain the imported MP14_SUBZONE_WEB_PL shapefile is called *mpsz* and it is a simple feature object. The geometry type is *multipolygon*. it is also important to note that mpsz simple feature object does not have EPSG information.

### **4.2 Updating CRS information**

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

After transforming the projection metadata, you can verify the projection of the newly transformed *mpsz_svy21* by using `st_crs()` of **sf** package.

The code chunk below will be used to verify the newly transformed *mpsz_svy21*.

```{r}
st_crs(mpsz_svy21)
```

Notice that the EPSG: is indicated as *3414* now.

To reveal the extent of *mpsz_svy21,* use `st_bbox()` of sf package.

```{r}
st_bbox(mpsz_svy21)
```

## **5 Aspatial Data Wrangling**

### **5.1 Importing the aspatial data**

The *condo_resale_2015* is in csv file format. The codes chunk below uses `read_csv()` function of **readr** package to import *condo_resale_2015* into R as a tibble data frame called *condo_resale*.

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

After importing the data file into R, it is important for us to examine if the data file has been imported correctly.

The codes chunks below uses `glimpse()` to display the data structure.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE)
```

```{r}
head(condo_resale$LATITUDE)
```

Next, `summary()` of base R is used to display the summary statistics of *cond_resale* tibble data frame.

```{r}
summary(condo_resale)
```

### **5.2 Converting aspatial data frame into a sf object**

Currently, the *condo_resale* tibble data frame is aspatial. We will convert it to a **sf** object. The code chunk below converts condo_resale data frame into a simple feature data frame by using `st_as_sf()` of **sf** packages.

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

Next, `head()` is used to list the content of *condo_resale.sf* object.

```{r}
head(condo_resale.sf)
```

Notice that the output is in point feature data frame.

## **6 Exploratory Data Analysis (EDA)**

Statistical graphics functions of **ggplot2** package to perform EDA.

### **6.1 EDA using statistical graphics**

We can plot the distribution of *SELLING_PRICE* by using appropriate Exploratory Data Analysis (EDA) as shown in the code chunk below.

```{r}
ggplot(data=condo_resale.sf, aes(x=`SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

The figure above shows a right-skewed distribution, indicating that more condominium units were sold at relatively lower prices.

Statistically, a skewed distribution can be normalized using a log transformation. The code snippet below demonstrates how to create a new variable called `LOG_SELLING_PRICE` by applying a log transformation to the `SELLING_PRICE` variable using the `mutate()` function from the `dplyr` package.

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

Plotting the *LOG_SELLING_PRICE* using the code chunk below

```{r}
ggplot(data=condo_resale.sf, aes(x=`LOG_SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

Notice that the distribution is relatively less skewed after the transformation.

### **6.2 Multiple Histogram Plots distribution of variables**

Drawing a small multiple histograms (also known as trellis plot) by using `ggarrange()` of [**ggpubr**](https://cran.r-project.org/web/packages/ggpubr/index.html) package.

The code chunk below is used to create 12 histograms. Then, `ggarrange()` is used to organised these histogram into a 3 columns by 4 rows small multiple plot.

```{r}
AREA_SQM <- ggplot(data=condo_resale.sf, aes(x= `AREA_SQM`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

AGE <- ggplot(data=condo_resale.sf, aes(x= `AGE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_CHILDCARE`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, PROX_ELDERLYCARE, 
          PROX_URA_GROWTH_AREA, PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

### **6.3 Drawing Statistical Point Map**

Lastly, we want to reveal the geospatial distribution condominium resale prices in Singapore. The map will be prepared by using **tmap** package.

First, we will turn on the interactive mode of tmap by using the code chunk below.

```{r}
tmap_mode("view")
```

Next, the code chunks below is used to create an interactive point symbol map.

```{r}
#| eval: false
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

Notice that [`tm_dots()`](https://www.rdocumentation.org/packages/tmap/versions/2.2/topics/tm_symbols) is used instead of `tm_bubbles()`.

`set.zoom.limits` argument of `tm_view()` sets the minimum and maximum zoom level to 11 and 14 respectively.

Before moving on to the next section, the code below will be used to turn R display into `plot` mode.

```{r}
tmap_mode("plot")
```

## **7 Hedonic Pricing Modelling in R**

Build hedonic pricing models for condominium resale units using [`lm()`](https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/lm) of R base.

### **7.1 Simple Linear Regression Method**

First, we will build a simple linear regression model by using *SELLING_PRICE* as the dependent variable and *AREA_SQM* as the independent variable.

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf)
```

The `lm()` function returns an object of class "lm" or, for multiple responses, a class of c("mlm", "lm").

The `summary()` and `anova()` functions can be used to generate and display a summary and an analysis of variance table for the results. Generic accessor functions such as `coefficients`, `effects`, `fitted.values`, and `residuals` extract various useful features from the object returned by `lm()`.

```{r}
summary(condo.slr)
```

The output report indicates that the selling price can be explained using the following formula:

y=−258121.1+14719x1​

The R-squared value of 0.4518 reveals that the simple regression model accounts for approximately 45% of the variation in resale prices.

Given that the p-value is much smaller than 0.0001, we reject the null hypothesis that the mean is a good estimator of the selling price. This allows us to infer that the simple linear regression model above is a good estimator of the selling price.

The "Coefficients" section of the report shows that the p-values for both the intercept and the coefficient of `ARA_SQM` are smaller than 0.001. This indicates that we can reject the null hypothesis that B0 and B1 are equal to 0. As a result, we can infer that B0 and B1 are good parameter estimates.

To visualize the best-fit curve on a scatterplot, we can incorporate `lm()` as a method function in `ggplot`'s geometry, as demonstrated in the code chunk below.

```{r}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}

```

Figure above reveals that there are a few statistical outliers with relatively high selling prices.

### **7.2 Multiple Linear Regression Method**

#### 7.2.1 Visualising the relationships of the independent variables

Before building a multiple regression model, it is crucial to ensure that the independent variables are not highly correlated with each other. Using highly correlated independent variables in a regression model can compromise its quality, a phenomenon known as multicollinearity in statistics.

A correlation matrix is commonly used to visualize the relationships between independent variables. Besides the `pairs()` function in R, many packages support the display of a correlation matrix. In this section, we will use the `corrplot` package.

The code chunk below is used to plot a scatterplot matrix of the relationships between the independent variables in the `condo_resale` data frame.

```{r}
#| fig-height: 10
#| fig-width: 12
corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")
```

Matrix reordering is essential for uncovering hidden structures and patterns within the matrix. The `corrplot` package in R offers four reordering methods (parameter `order`): "AOE", "FPC", "hclust", and "alphabet". In the code chunk above, the "AOE" order is used. This method, suggested by Michael Friendly, orders the variables using the angular order of the eigenvectors.

From the scatterplot matrix, it is evident that `Freehold` is highly correlated with `LEASE_99YEAR`. Given this high correlation, it is prudent to include only one of these variables in the subsequent model building. Therefore, `LEASE_99YEAR` is excluded from the model.

### **7.3 Building a hedonic pricing model using multiple linear regression method**

The code chunk below use `lm()` to calibrate the multiple linear regression model.

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale.sf)
summary(condo.mlr)
```

### **7.4 Preparing Publication Quality Table: olsrr method**

With reference to the report above, it is clear that not all the independent variables are statistically significant. We will revised the model by removing those variables which are not statistically significant.

Now, we are ready to calibrate the revised model by using the code chunk below.

```{r}
condo.mlr1 <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                   PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                   PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK + 
                   PROX_PRIMARY_SCH + PROX_SHOPPING_MALL    + PROX_BUS_STOP + 
                   NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                 data=condo_resale.sf)
ols_regress(condo.mlr1)
```

### **7.5 Preparing Publication Quality Table: gtsummary method**

The [**gtsummary**](https://www.danieldsjoberg.com/gtsummary/index.html) package provides an elegant and flexible way to create publication-ready summary tables in R.

In the code chunk below, [`tbl_regression()`](https://www.danieldsjoberg.com/gtsummary/reference/tbl_regression.html) is used to create a well formatted regression report.

```{r}
tbl_regression(condo.mlr1, intercept = TRUE)
```

With gtsummary package, model statistics can be included in the report by either appending them to the report table by using [`add_glance_table()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) or adding as a table source note by using [`add_glance_source_note()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) as shown in the code chunk below.

```{r}
tbl_regression(condo.mlr1, 
               intercept = TRUE) %>% 
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, adj.r.squared, 
                AIC, statistic,
                p.value, sigma))
```

```{r}

```

```{r}

```

#### 7.5.1 Checking for multicollinearity

To introduce a fantastic R package specially programmed for performing OLS regression called [**olsrr**](https://olsrr.rsquaredacademy.com/). It provides a collection of very useful methods for building better multiple linear regression models:

-   comprehensive regression output

-   residual diagnostics

-   measures of influence

-   heteroskedasticity tests

-   collinearity diagnostics

-   model fit assessment

-   variable contribution assessment

-   variable selection procedures

In the code chunk below, the [`ols_vif_tol()`](https://olsrr.rsquaredacademy.com/reference/ols_coll_diag.html) of **olsrr** package is used to test if there are sign of multicollinearity.

```{r}
ols_vif_tol(condo.mlr1)
```

Since the VIF of the independent variables are less than 10. We can safely conclude that there are no sign of multicollinearity among the independent variables.

#### 7.5.2 Test for Non-Linearity

In multiple linear regression, it is important for us to test the assumption that linearity and additivity of the relationship between dependent and independent variables.

In the code chunk below, the [`ols_plot_resid_fit()`](https://olsrr.rsquaredacademy.com/reference/ols_plot_resid_fit.html) of **olsrr** package is used to perform linearity assumption test.

```{r}
ols_plot_resid_fit(condo.mlr1)
```

The figure above reveals that most of the data poitns are scattered around the 0 line, hence we can safely conclude that the relationships between the dependent variable and independent variables are linear.

#### 7.5.3 Test for Normality Assumption

Lastly, the code chunk below uses [`ols_plot_resid_hist()`](https://olsrr.rsquaredacademy.com/reference/ols_plot_resid_hist.html) of *olsrr* package to perform normality assumption test.

```{r}
ols_plot_resid_hist(condo.mlr1)
```

The figure reveals that the residual of the multiple linear regression model (i.e. condo.mlr1) resembles normal distribution.

If you prefer formal statistical test methods, the [`ols_test_normality()`](https://olsrr.rsquaredacademy.com/reference/ols_test_normality.html) of **olsrr** package can be used as shown in the code chunk below.

```{r}
#| warning: false
ols_test_normality(condo.mlr1)
```

The summary table above reveals that the p-values of the four tests are way smaller than the alpha value of 0.05. Hence we will reject the null hypothesis and infer that there is statistical evidence that the residual are not normally distributed.

#### 7.5.4 Testing for Spatial Autocorrelation

The hedonic model we try to build are using geographically referenced attributes, hence it is also important for us to visual the residual of the hedonic pricing model.

In order to perform spatial autocorrelation test, we need to convert *condo_resale.sf* from sf data frame into a **SpatialPointsDataFrame**.

First, we will export the residual of the hedonic pricing model and save it as a data frame.

```{r}
mlr.output <- as.data.frame(condo.mlr1$residuals)
```

Next, we will join the newly created data frame with *condo_resale.sf* object.

```{r}
condo_resale.res.sf <- cbind(condo_resale.sf, 
                        condo.mlr1$residuals) %>%
rename(`MLR_RES` = `condo.mlr1.residuals`)
```

Next, we will convert *condo_resale.res.sf* from simple feature object into a SpatialPointsDataFrame because spdep package can only process sp conformed spatial data objects.

The code chunk below will be used to perform the data conversion process.

```{r}
condo_resale.sp <- as_Spatial(condo_resale.res.sf)
condo_resale.sp
```

Next, we will use **tmap** package to display the distribution of the residuals on an interactive map.

The code churn below will turn on the interactive mode of tmap.

```{r}
tmap_mode("view")
```

The code chunks below is used to create an interactive point symbol map.

```{r}
tm_shape(mpsz_svy21)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(condo_resale.res.sf) +  
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

Switch back to “plot” mode before continue.

```{r}
tmap_mode("plot")
```

The figure above reveal that there is sign of spatial autocorrelation.

To proof that our observation is indeed true, the Moran’s I test will be performed

First, we will compute the distance-based weight matrix by using [`dnearneigh()`](https://r-spatial.github.io/spdep/reference/dnearneigh.html) function of **spdep**.

```{r}
#| warning: false
nb <- dnearneigh(coordinates(condo_resale.sp), 0, 1500, longlat = FALSE)
summary(nb)
```

Next, [`nb2listw()`](https://r-spatial.github.io/spdep/reference/nb2listw.html) of **spdep** packge will be used to convert the output neighbours lists (i.e. nb) into a spatial weights.

```{r}
nb_lw <- nb2listw(nb, style = 'W')
summary(nb_lw)
```

Next, [`lm.morantest()`](https://r-spatial.github.io/spdep/reference/lm.morantest.html) of **spdep** package will be used to perform Moran’s I test for residual spatial autocorrelation

```{r}
lm.morantest(condo.mlr1, nb_lw)
```

The Global Moran’s I test for residual spatial autocorrelation shows that it’s p-value is less than 0.00000000000000022 which is less than the alpha value of 0.05. Hence, we will reject the null hypothesis that the residuals are randomly distributed.

Since the Observed Global Moran I = 0.1424418 which is greater than 0, we can infer than the residuals resemble cluster distribution.

### **8.1 Building Fixed Bandwidth GWR Model**

#### 8.1.1 Computing fixed bandwith

In the code chunk below `bw.gwr()` of GWModel package is used to determine the optimal fixed bandwidth to use in the model. Notice that the argument ***adaptive*** is set to **FALSE** indicates that we are interested to compute the fixed bandwidth.

There are two possible approaches can be uused to determine the stopping rule, they are: CV cross-validation approach and AIC corrected (AICc) approach. We define the stopping rule using ***approach*** argeement.

```{r}
bw.fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                     PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                     PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                     PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                     FAMILY_FRIENDLY + FREEHOLD, 
                   data=condo_resale.sp, 
                   approach="CV", 
                   kernel="gaussian", 
                   adaptive=FALSE, 
                   longlat=FALSE)
```

The result shows that the recommended bandwidth is 971.3405 metres.

It is in metres due to it reflecting the units used for the coordinate system in spatial data.

#### 8.1.2 GWModel method - fixed bandwith

Now we can use the code chunk below to calibrate the gwr model using fixed bandwidth and gaussian kernel.

```{r}
gwr.fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                         PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                         PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                         PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                         FAMILY_FRIENDLY + FREEHOLD, 
                       data=condo_resale.sp, 
                       bw=bw.fixed, 
                       kernel = 'gaussian', 
                       longlat = FALSE)
```

The output is saved in a list of class “gwrm”. The code below can be used to display the model output.

```{r}
gwr.fixed
```

The report shows that the AICc of the gwr is 42263.61 which is significantly smaller than the globel multiple linear regression model of 42967.1.

### **8.2 Building Adaptive Bandwidth GWR Model**

In this section, we will calibrate the gwr-based hedonic pricing model by using adaptive bandwidth approach.

#### 8.2.1 Computing the adaptive bandwidth

Similar to the earlier section, we will first use `bw.gwr()` to determine the recommended data point to use.

The code chunk used look very similar to the one used to compute the fixed bandwidth except the `adaptive` argument has changed to **TRUE**.

```{r}
bw.adaptive <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
                        PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    + 
                        PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                        PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + PROX_BUS_STOP + 
                        NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                      data=condo_resale.sp, 
                      approach="CV", 
                      kernel="gaussian", 
                      adaptive=TRUE, 
                      longlat=FALSE)
```

The result shows that the 30 is the recommended data points to be used.

#### 8.2 Constructing the adaptive bandwidth gwr model

Now, we can go ahead to calibrate the gwr-based hedonic pricing model by using adaptive bandwidth and gaussian kernel as shown in the code chunk below.

```{r}
gwr.adaptive <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                            PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + 
                            PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                            PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + 
                            NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                          data=condo_resale.sp, bw=bw.adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE, 
                          longlat = FALSE)
```

The code below can be used to display the model output.

```{r}
gwr.adaptive
```

The report shows that the AICc the adaptive distance gwr is 41982.22 which is even smaller than the AICc of the fixed distance gwr of 42263.61.

### **8.3 Visualising GWR Output**

In addition to regression residuals, the output feature class table contains fields for observed and predicted y values, condition number (cond), Local R2, residuals, as well as coefficients and standard errors for the explanatory variables:

-   **Condition Number**: This diagnostic assesses local collinearity. When strong local collinearity exists, the results may become unstable. Condition numbers greater than 30 may indicate unreliable results.

-   **Local R2**: These values range from 0.0 to 1.0 and reflect how well the local regression model fits the observed y values. Very low R2 values suggest that the local model performs poorly. Mapping the Local R2 values can help identify areas where GWR predicts well versus those where it does not, potentially highlighting important variables that are missing from the regression model.

-   **Predicted**: These represent the estimated (or fitted) y values computed by GWR.

-   **Residuals**: Residual values are obtained by subtracting the fitted y values from the observed y values. Standardized residuals have a mean of zero and a standard deviation of one. A cold-to-hot rendered map of standardized residuals can be created using these values.

-   **Coefficient Standard Error**: These values indicate the reliability of each coefficient estimate. Higher confidence in these estimates is associated with smaller standard errors relative to the actual coefficient values. Large standard errors may suggest issues with local collinearity.

All of this information is stored in a SpatialPointsDataFrame or SpatialPolygonsDataFrame object, which is integrated with fit.points, GWR coefficient estimates, y values, predicted values, coefficient standard errors, and t-values within the “data” slot of an object called **SDF** in the output list.

### **8.4 Converting SDF into *sf* data.frame**

To visualise the fields in **SDF**, we need to first covert it into **sf** data.frame by using the code chunk below.

```{r}
condo_resale.sf.adaptive <- st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
```

```{r}
condo_resale.sf.adaptive.svy21 <- st_transform(condo_resale.sf.adaptive, 3414)
condo_resale.sf.adaptive.svy21  
```

```{r}
gwr.adaptive.output <- as.data.frame(gwr.adaptive$SDF)
condo_resale.sf.adaptive <- cbind(condo_resale.res.sf, as.matrix(gwr.adaptive.output))
glimpse(condo_resale.sf.adaptive)
```

### **8.5 Visualising local R2**

The code chunks below is used to create an interactive point symbol map.

```{r}
tmap_mode("view")
tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
#| echo: false
tmap_mode("plot")
```

### **8.6 Visualising coefficient estimates**

The code chunks below is used to create an interactive point symbol map.

```{r}
tmap_mode("view")
AREA_SQM_SE <- tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_SE",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

AREA_SQM_TV <- tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "AREA_SQM_TV",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_arrange(AREA_SQM_SE, AREA_SQM_TV, 
             asp=1, ncol=2,
             sync = TRUE)
```

```{r}
#| echo: false
tmap_mode("plot")
```

#### 8.6.1 By URA Plannign Region

```{r}
tm_shape(mpsz_svy21[mpsz_svy21$REGION_N=="CENTRAL REGION", ])+
  tm_polygons()+
tm_shape(condo_resale.sf.adaptive) + 
  tm_bubbles(col = "Local_R2",
           size = 0.15,
           border.col = "gray60",
           border.lwd = 1)
```

## **13.10 Reference**

Gollini I, Lu B, Charlton M, Brunsdon C, Harris P (2015) “GWmodel: an R Package for exploring Spatial Heterogeneity using Geographically Weighted Models”. *Journal of Statistical Software*, 63(17):1-50, http://www.jstatsoft.org/v63/i17/

Lu B, Harris P, Charlton M, Brunsdon C (2014) “The GWmodel R Package: further topics for exploring Spatial Heterogeneity using GeographicallyWeighted Models”. *Geo-spatial Information Science* 17(2): 85-101, http://www.tandfonline.com/doi/abs/10.1080/1009502.2014.917453

