---
title: "In-class Exercise 2"
author: "Ryan Peh"
date: "August 26, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

Required packages are installed using the code chunk below.

```{r}
pacman::p_load(sf, tidyverse)
```

This import shapefile of mpsz 2014

```{r}
mpsz2014 = st_read(dsn = "data", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

Check if class is data.frame

```{r}
class(mpsz2014)
```

Importing kml file of mpsz2014

```{r}
st_write(mpsz2014, "data/MasterPlan2014SubzoneBoundaryWebKML.kml", delete_dsn = TRUE)
```

```{r}
mpsz2014_kml = st_read(dsn = "data/MasterPlan2014SubzoneBoundaryWebKML.kml")
```

This import shapefile of mpsz 2019

```{r}
mpsz2019 = st_read(dsn = "data", 
                  layer = "MPSZ-2019")
```

Check if class is data.frame

```{r}
class(mpsz2019)
```

Importing kml file of mpsz2019

```{r}
mpsz2019_kml = st_read(dsn = "data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

Handling Coordinate System

check coordinate system

```{r}
st_crs(mpsz2019)
```

Transform coordinate system

```{r}
mpsz2014 <- st_read(dsn = "data", 
                  layer = "MP14_SUBZONE_WEB_PL") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz2019 <- st_read(dsn = "data", 
                  layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

working with pop data

```{r}
popdata <- read.csv("data/respopagesextod2023.csv")
```

data wrangling

```{r}
popdata2023 <- popdata %>% 
  group_by(PA, SZ, AG) %>% 
  summarise(`POP`=sum(`Pop`)) %>% 
  ungroup() %>%
  pivot_wider(names_from = AG, 
              values_from = POP)

colnames(popdata2023)
```

data processing

```{r}
popdata2023 <- popdata2023 %>%
  mutate(`YOUNG` = rowSums(.[3:6]) + rowSums(.[14])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:13])+rowSums(.[15])) %>%
  mutate(`AGED` = rowSums(.[16:21])) %>%
  mutate(`TOTAL` = rowSums(.[3:21])) %>%
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`, `DEPENDENCY`)

```

```{r}
popdata2023 <- popdata2023 %>%
  mutate_at(.vars = vars(PA, SZ), 
            .funs = list(toupper)) 

```

```{r}
mpsz_pop2023 <- left_join(mpsz2019, popdata2023, 
                          by = c("SUBZONE_N" = "SZ"))

mpsz_pop2023

```
