---
title: "Take-home_Ex01"
author: "Ryan Peh"
date: "September 7 30, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---

## **1 Overview**

### **1.1 Objective**

To apply spatial point patterns analysis methods to discover the spatial and spatio-temporal distribution of armed conflict in Myanmar.

### 1.2 Background

## **2 Setup**

### **2.1 Data**

-   For the purpose of this assignment, armed conflict data of Myanmar between 2021-2024 from [Armed Conflict Location & Event Data (ACLED)](https://acleddata.com/), an independent, impartial, international non-profit organization collecting data on violent conflict and protest in all countries and territories in the world, should be used.

    In terms of event types, students should focus on at least four main event types, namely: Battles, Explosion/Remote violence, Strategic developments, and Violence against civilians.

    In terms of study period, students should focus on quarterly armed conflict events from January 2021 until June 2024.

-   [Geospatial data](http://geonode.themimu.info/layers/?limit=100&offset=0) on [Myanmar Information Management Unit, MIMU](http://www.themimu.info/)

## **3 The Task**

The specific tasks of this take-home exercise are as follows:

-   Using appropriate function of **sf** and **tidyverse** packages, import and transform the downloaded armed conflict data and administrative boundary data into sf tibble data.frames.

-   Using the geospatial data sets prepared, derive quarterly KDE layers.

-   Using the geospatial data sets prepared, perform 2nd-Order Spatial Point Patterns Analysis.

-   Using the geospatial data sets prepared, derive quarterly spatio-temporal KDE layers.

-   Using the geospatial data sets prepared, perform 2nd-Order Spatio-temporal Point Patterns Analysis.

-   Using appropriate **tmap** functions, display the KDE and Spatio-temporal KDE layers on openstreetmap of Myanmar.

-   Describe the spatial patterns revealed by the KDE and Spatio-temporal KDE maps.

```{r}
pacman::p_load(tmap, sf, tidyverse, raster, spatstat, sp)
```

```{r}
myanmar <- st_read(dsn = "data/raw/geospatial", 
                 layer = "mmr_polbnda_adm1_250k_mimu_1")
```

```{r}
myanmar
```

```{r}
myanmar_conflict <- read_csv("data/raw/ACLED_Myanmar_2021_2024.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
                       crs = 4326)  %>%
  st_transform(crs = 32646)
```

```{r}
tm_shape(myanmar) +
  tm_polygons()
```

```{r}

```

```{r}

```